Figures for Paper

### 2021-09-17 paper_figures.Rmd
update the index for rendering

only_short_plotter %>% 
    ggplot(aes(x = detection_name,
               y = inclusion_reads * 10^6,
               fill ="red")) +
    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single",width = 1),outlier.colour = NA) +
    geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+
    ggplot2::facet_wrap(vars(tissue_clean),scales = "free_x",nrow = 3) +
    ylab("UNC13A cryptic \n reads per million") +
    theme(text = element_text(size = 12)) +
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    theme(text = element_text(size = 18))  +
    theme(legend.position = "bottom") + 
    stat_compare_means(comparisons = comps)
```

@@

```
## Detection by genotype
```{r}

clean_data_table %>%
  filter(!(rs12973192 == "C/G" & rs12608932 == "C/C")) %>%
  mutate(snp_status = case_when(rs12973192 == "C/C" ~ "healthy/healthy",
                                rs12973192 == "C/G" ~ "healthy/risk",
                                rs12973192 == "G/G" ~ "risk/risk")) %>% 
  filter(disease_tissue == T) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%

@@

```{r str and log2 graph}
str_genotype = janitor::clean_names(fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/rs56041637.genotypes.377samples.txt")) %>% 
  dplyr::rename(presumed_unique_id = number_sample_id)

@@

ipsc_de[gene_name %in% als_ftd_genomics][order(log2FoldChange)][,.(log2FoldChange,padj,gene_name,gene_expresion)] %>% mutate(has_cryptic = gene_name %in% ipsc_splicing[cryptic_junction == T,gene_name]) %>% unique() %>% arrange(log2FoldChange) %>% clipr::write_clip()
```

## Linear model on UNC13A CE PSI with covariates
```{r the_silly_linear}
# library size 
#llumina HiSeq 2500 (125 bp paired end) or an Illumina NovaSeq (100 bp paired end). 
full_cov = clean_data_table %>%
    filter(tissue_clean %in% c("Frontal_Cortex", 
                               "Lumbar_Spinal_Cord", 
                               "Cervical_Spinal_Cord", 
                               "Motor_Cortex", 
                               "Temporal_Cortex")) %>% 
    mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
    mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
    filter(sequencing_platform != "") %>% 
    left_join(cell_types %>% 
                  dplyr::select(cell,deconv,sample) %>% 
                  unique() %>% 
                  pivot_wider(id_cols = "sample",names_from = "cell",values_from = "deconv"), by = 'sample') %>%
    filter(!is.na(disease_group2)) %>% 
    unique() %>% 
    filter(disease_group2 %in% c("ALS-TDP","FTLD-TDP")) %>% 
    dplyr::select(c("unc13a_cryptic_leaf_psi",
                    "sample",  "tissue_clean",  
                    "rin",  "sex", 
                    "age",  "mutations", 
                    "sequencing_platform", 
                    "number_g_alleles", 
                    "stmn_2_cryptic_psi_leaf", "oligodendrocytes", 
                    "astrocytes", "neurons", "endothelial", "microglia")) %>% 
    column_to_rownames('sample')

@@

fit %>% broom::tidy() %>% fwrite(paste0(here::here(),"/covariate_unc13a_ce.csv"))
full_cov %>% 
    group_by(tissue_clean) %>% 
    nest() %>% 
    mutate(lm_obj = map(data, ~lm(unc13a_cryptic_leaf_psi ~., data = .x))) %>% 
    mutate(broomed = map(lm_obj, broom::tidy)) %>% 
    dplyr::select(tissue_clean,broomed) %>% unnest() %>% filter(p.value < 0.05) %>% View()

nested %>% 
    pluck("lm_obj", 1) %>% broom::tidy()
unc13a_data_fisher = unc13a_sequencing_platform %>% 
    filter(tissue_clean != "Thoracic_Spinal_Cord") %>% 
    group_by(disease_group2,tissue_clean) %>% 
    nest() %>% 
    mutate(fisher_pvalue = map_dbl(data, quick_fisher))

@@

guide_strength %>% 
  melt() %>% 
  filter(`cell line` == "i11w_repeat") %>% 
   mutate(condition = fct_relevel(condition,
                              "control","tdp43_mild")) %>%
    ggbarplot(,
              x = "condition",
              add = c("mean_se","jitter"),
              y = "value",
              fill = 'variable',
              color = 'variable',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
      scale_color_manual(
        values = c("#1C2617","#262114","#262114")
    ) +
        scale_fill_manual(
        values = c("#1576b8","#eb7f3f","#975452")
    ) +
    scale_y_continuous(labels = scales::percent) +
  ggtitle("WTC11") +
          axis.line = element_line(colour = 'black', size = 1.5),
        axis.ticks = element_line(colour = "black", size = 2))
```

@@

agg=aggregate(Value~Treatment*Gene*Line_Group*Color_Group, data=Protein, FUN="mean")
standard_error <- function(x) sd(x) / sqrt(length(x))
agg$se=aggregate(Value~Treatment*Gene*Line_Group*Color_Group, data=Protein, FUN="standard_error")$Value

@@

agg = agg %>% 
  filter(Gene != "13A-RNA") %>% 
  mutate(Line_Group = gsub("Correct_RNA","RNA",Line_Group))

>>>>

only_short_plotter %>% 
    ggplot(aes(x = detection_name,
               y = inclusion_reads * 10^6,
               fill ="red")) +
    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single",width = 1),outlier.colour = NA) +
    geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+
    ggplot2::facet_wrap(vars(tissue_clean),scales = "free_x",nrow = 3) +
    ylab("UNC13A cryptic \n reads per million") +
    theme(text = element_text(size = 12)) +
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    theme(text = element_text(size = 18))  +
    theme(legend.position = "bottom") 
```

@@

@@

```{r str and log2 graph}
str_genotype = janitor::clean_names(fread(file.path(here::here(),"data/rs56041637.genotypes.377samples.txt"))) %>% 
  dplyr::rename(participant_id = number_sample_id)

@@

ipsc_de[gene_name %in% als_ftd_genomics][order(log2FoldChange)][,.(log2FoldChange,padj,gene_name,gene_expresion)] %>% mutate(has_cryptic = gene_name %in% ipsc_splicing[cryptic_junction == T,gene_name]) %>% unique() %>% arrange(log2FoldChange) %>% knitr::kable()
```

# Linear model on UNC13A CE PSI with covariates
```{r the_silly_linear}
# library size 
#llumina HiSeq 2500 (125 bp paired end) or an Illumina NovaSeq (100 bp paired end). 
full_cov = clean_data_table %>%
    filter(tissue_clean %in% c("Frontal_Cortex", 
                               "Lumbar_Spinal_Cord", 
                               "Cervical_Spinal_Cord", 
                               "Motor_Cortex", 
                               "Temporal_Cortex")) %>% 
    mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
    mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
    filter(sequencing_platform != "") %>% 
    left_join(cell_types %>% 
                  dplyr::select(cell,deconv,sample) %>% 
                  unique() %>% 
                  pivot_wider(id_cols = "sample",names_from = "cell",values_from = "deconv"), by = 'sample') %>%
    filter(!is.na(disease_group2)) %>% 
    unique() %>% 
    filter(disease_group2 %in% c("ALS-TDP","FTLD-TDP")) %>% 
    dplyr::select(c("unc13a_cryptic_leaf_psi",
                    "sample",  "tissue_clean",  
                    "rin",  "sex", 
                    "age",  "mutations", 
                    "sequencing_platform", 
                    "number_g_alleles", 
                    "stmn_2_cryptic_psi_leaf", "oligodendrocytes", 
                    "astrocytes", "neurons", "endothelial", "microglia")) %>% 
    column_to_rownames('sample')

@@

fit %>% broom::tidy() %>% fwrite(paste0(here::here(),"/covariate_unc13a_ce.csv"))
full_cov %>% 
    group_by(tissue_clean) %>% 
    nest() %>% 
    mutate(lm_obj = map(data, ~lm(unc13a_cryptic_leaf_psi ~., data = .x))) %>% 
    mutate(broomed = map(lm_obj, broom::tidy)) %>% 
    dplyr::select(tissue_clean,broomed) %>% 
    unnest() %>% 
    filter(p.value < 0.05) %>% knitr::kable()

```{r detection_seq_platform}
unc13a_data_fisher = unc13a_sequencing_platform %>% 
    filter(tissue_clean != "Thoracic_Spinal_Cord") %>% 
    group_by(disease_group2,tissue_clean) %>% 
    nest() %>% 
    mutate(fisher_pvalue = map_dbl(data, quick_fisher))

@@

guide_strength %>% 
  melt() %>% 
  filter(`cell line` == "i11w_repeat") %>% 
   mutate(condition = fct_relevel(condition,
                              "control","tdp43_mild")) %>%
    ggbarplot(,
              x = "condition",
              add = c("mean_se","jitter"),
              y = "value",
              fill = 'variable',
              color = 'variable',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
      scale_color_manual(
        values = c("#1C2617","#262114","#262114")
    ) +
        scale_fill_manual(
        values = c("#1576b8","#eb7f3f","#975452")
    ) +
    scale_y_continuous(labels = scales::percent) +
  ggtitle("WTC11") +
           theme(axis.line = element_line(colour = 'black', size = 1.5),
        axis.ticks = element_line(colour = "black", size = 2))
```

@@

agg=aggregate(Value~Treatment*Gene*Line_Group*Color_Group, data=Protein, FUN="mean")
std <- function(x) sd(x) / sqrt(length(x))

tmp = Protein %>% 
  group_by(Treatment,Gene,Line_Group,Color_Group) %>% 
  summarise(se = std(Value))

@@

agg = agg %>% 
  filter(Gene != "13A-RNA") %>% 
  mutate(Line_Group = gsub("Correct_RNA","RNA",Line_Group)) %>% 
  left_join(tmp)
